The Talking Heads Experiment

### 2015-04-10 chap9/chap9.tex
fixes
quotation marks for chapters 8 to 10

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.75\textwidth]{chap9/figs/cambridge-view}}
\caption{\label{fig:sideview} 
View through the street window inside the Kettle's Yard gallery in Cambridge. The cameras are located before 
the towers. The geometric figures are attached to the wall and explanations of the experiment are located on the right.}
\end{figure}

The N01SE exhibition was curated by Adam Lowe (see \figref{fig:lowe}) and Simon Schaffer. 
Adam Lowe is an artist and technologist. He is currently the director of Factum arte (Madrid) which is specialised 
in making life-like replicas of paintings, sculptures and archeological objects, using laser-scanners and 3d printers. 
Recent realisations include the sculptures of Giambattista Piranesi \citep{Lowe:2010}, the 
paintings of Caravaggio and the tomb of Tutankhamun. 
Simon Schaffer is a science historian, professor at the University of Cambridge at the department of History and Philosophy 
of Science. He wrote extensively about the historical developments in scientific research \citep{Schaffer:2011}
and animated television and radio programs, including {\itshape Light Fantastic} (BBC4). 
\begin{figure}[htbp]
  \centerline{\includegraphics[width=.70\textwidth]{chap9/figs/lowe}}
\caption{\label{fig:lowe} 
Adam Lowe (curator of N01SE) during installation of the Talking Heads experiment at the Wellcome Gallery in London. 
The backwall contained the figures. On the right wall, posters were displayed explaining the experiment.}
\end{figure}

@@

N01SE celebrates the essential excess from which information is drawn. It probes many different ways of seeing and 
being in the world. Chances are your own sense of order is already someone elses N01SE.}
\end{quotation}
%\begin{figure}[htbp]
%  \centerline{\includegraphics[width=.95\textwidth]{chap9/figs/latour}}
%\caption{\label{latour} 
%Luc Steels Discusses The N01SE exhibition with Bruno Latour who intensively followed the experiments already from the first series at the Laboratorium exhibition in Antwerp. His comments significantly improved the initial designs.}
%\end{figure}

@@

To play the game, the Speaker segments the scene and performs pattern recognition to extract features - area, shape, colour, position - about the segments selected. The Speaker then conceptualises the focus element ``topic" as distinct from those of other segments - be it the largest or the furthest to the left or the green one. Next the Speaker verbalises this conceptualisation using descriptive words selected from its lexicon. The Hearer works the other way around; it queries the words transmitted by the Speaker and applies the resultant meanings back to the scene to find what topic the Speaker intended.

@@

The installation consists of two computer-controlled robotic camera heads that capture images from `scenes' in front of them consisting of colored geometric shapes pasted on a magnetic whiteboard. The configurations on the board can be changed at any time, making the robots' world unpredictable and open.

@@

\begin{itemize} 
\item Two embodied grounded agents always see the situation from different points of view so that they capture different images. Consequently they may have divergent perceptions and hence great difficulty to arrive at a successful game. 
\item The word(s) transmitted may not be accurately produced or received. For example, one agent may produce ``wabaku'' but the other agent may hear ``mabaku''. This introduces noise in the signal itself and hence possibly confusion among the agents. 
\item A scene can usually be conceptualised in many different ways, so that there is seldom certainty among the agents whether they share the same meaning. This causes great difficulties in learning the meaning of unknown words. 
\item The lexicons and conceptual repertoires are never exactly the same as each agent develops them autonomously. This generates in additional sources of confusion.
\end{itemize}

@@

Another factor in language evolution is due to changes in the environment. Thus two alternative meanings for a word may be compatible for a while but are then disambiguated when a series of scenes arises in which the two meanings are no longer both applicable. For example, all objects may be both green and small and therefore there may be a word ``sesubipu'' which may mean both, until a clear situation arises where a green object is no longer the smallest and a misunderstanding arises.

Semiotic evolution is continually present. Different meanings for a particular word will emerge over over a large number of language games. During specific periods, different words dominate. The word ``droite'' (originally introduced by a French speaker) gains the dominant meaning ``to the right'', then shifts to ``at the bottom'', and then to ``very much to the right''. Particularly the words introduced by humans have a tendency to undergo this kind of strong evolution because human users do not know which meanings their agents employ.

@@

Each agent has been given a mechanism to `grow' new distinctions by expanding discrimination trees. Each tree discretises one sensory dimension. For example, there is a tree for the area of a segment (scale with respect to the image) which divides the range of possible values into two discrete regions, which would be named in English ``small'' and ``large''Other trees focus on position (left versus right or top versus bottom), shape (rectangular or oval), color, etc. Trees can go as deep as necessary to carve out smaller and smaller subregions of a continuous space.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.60\textwidth]{chap9/figs/NOISE-programmaboek}}
\caption{\label{fig:program} 
Catalogue cover and poster of the N01SE exhibition at the Wellcome Gallery in London.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.55\textwidth]{chap9/figs/london-heads} 
\includegraphics[width=.44\textwidth]{chap9/figs/screen}}
\caption{\label{fig:london-heads} 
Installation at the Wellcome Gallery in London. Left: Talking Heads cameras oriented towards the wall on which 
geometric figures were pasted. Right: Projection of interaction during ongoing experiment. A game just failed and 
the speaker says ``no''.} 
\end{figure}

@@

\noindent
The installation consists of two robotic heads. These are steerable cameras controlled by a computer that hosts the 
architecture and knowledge state of each agent. The robots capture images from scenes in front of them. The scenes consist 
of coloured geometrical shapes pasted on a magnetic white board. The configuration on the board can be changed at any 
time, making the robots' world unpredictable and open. The robot infrastructure is connected to the Internet, so 
that an agent may dematerialize from a body and travel over the internet to another body in which it can re-materialize.
Two robotic structures will be active as part of the N01SE exhibition: one at Kettles'Yard in Cambridge and another 
one at the Wellcome Trust Two10 Gallery in London.

@@

\noindent
What is most remarkable about this experiment is that the robotic agents do not come with 
pre-programmed ways of conceptualizing reality but have to develop their own concepts. Each agent has been 
given a mechanism to `grow' new distinctions by expanding discrimination trees. Each tree discretizes one sensory 
dimension. For example, there is a tree for the area of a segment which divides the range of possible values
into two discrete regions, which would be named in English `small' and `large'. Other trees focus on position 
(left versus right or top versus bottom), shape (rectangular or oval), colour, etc.

@@

\noindent
The creation of a shared language by a group of autonomous distributed agents is extraordinarily difficult 
because there are many sources of noise, in the form of disturbances that cause incoherence between the agents: 
\begin{itemize} 
\item Two embodied grounded agents always see the situation from different points of view so that they capture 
different images. Consequently they may have divergent perceptions and hence great difficulty to arrive at a 
successful game. 
\item The word(s) transmitted may not be accurately produced or received. For example, one agent may produce 
'wabaku' but the other agent may hear `mabakau'. This introduces noise in the signal itself and hence 
possibly confusion among the agents. 
\item A scene can usually be conceptualized in many different ways, so that there is seldom certainty among 
the agents whether they share the same meaning. This causes great difficulties in learning 
the meaning of unknown words. 
\item The lexicons and conceptual repertoires are never exactly the same due to the fact that each agent
develops them autonomously. This brings in additional sources of confusion. 
\end{itemize}
Any theory claiming to explain the origins of word meaning must confront the handling of 
noise head on.

@@

\noindent
Noise plays yet another role, namely as a motor of language evolution. Indeed natural lexicons evolve - 
even if there are already perfectly good words in a language. Noise on the word form causes changes 
in the form which propagate in the remainder of the population. Misunderstandings may destabilize a
word and cause its meaning to shift.

@@

Many aspects of this episode are remarkable, not in the least that those responsable for an 
exhibition (i.c. Denna Jones (i.e. phlox) and `fish') were bent on creating a wave of negative reactions and 
destructions against one of the exhibition pieces entrusted to them. Apparently this behaviour was triggered through
a conflict 
with Adam Lowe, curator of the N01SE exhibition, but the team behind the Talking Heads experiment had nothing to do 
with this. Another remarkable fact is that the hacker group 
became aggressive as soon as they felt they were no longer able to have control the way they wanted to, specifically 
to introduce disrespectful language or to subvert agent scripts to circumvent the central scheduler so that they 
could send their agents in priority to the sites of their choosing.
This style of behaviour is a personality trait which is commonly recognised as characteristic for 
hackers, including by members of the hacker community themselves:\footnote{
Raymond, E. (2013) The new hackers dictionary. Available as http://www.catb.org/jargon/html/index.html.}

\begin{quotation}
"Hackers have relatively little ability to identify emotionally with other people. ... Unsurprisingly, hackers also tend towards self-absorption, intellectual arrogance, and impatience with people and tasks perceived to be wasting their time. Because of their passionate embrace of (what they consider to be) the Right Thing, hackers can be unfortunately intolerant and bigoted on technical issues, in marked contrast to their general spirit of camaraderie and tolerance of alternative viewpoints otherwise. ... As a result of all the above traits, many hackers have difficulty maintaining stable relationships. At worst, they can produce the classic geek: withdrawn, relationally incompetent, sexually frustrated, and desperately unhappy when not submerged in his or her craft."
\end{quotation}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap9/figs/th4}}
\caption{\label{fig:palais} 
Talking Heads installation at the Palais de la D\'{e}couverte. It featured new fancy structures that 
made the installation more attractive visually. Daily explanations were given to visitors 
by the staff of the Palais de la D\'{e}couverte.} 
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.60\textwidth]{chap9/figs/road}}
\caption{\label{fig:road} 
Portable installation with two pan-tilt cameras and a portable computer that was able to run the TH software.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.90\textwidth]{chap9/figs/Lausanne}}
\caption{\label{fig:lausanne} 
Live installation and demonstration of the Talking Heads experiment at the European Conference on Artificial 
Life in 1999.}
\end{figure}

2. The {\itshape Neuer Aachener Kunstverein} in Aachen (Germany) in collaboration with the 
RWTH (the technical university of Aachen) organised a general exhibition on modelling called 
Modell-Modell. Within this framework, artist Anne-Mie van Kerckhoven invited me to cooperate in a laboratory on 
language and colour called ``Cyberlabor - Chromosophy''. As part of this laboratory, the portable Talking Heads 
experiment was installed and ran from 5 May to 16 June 2000. It featured not only the portable Talking Heads, 
which was demonstrated live, but also posters, talks about the project and additional art works.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap9/figs/mich-gallery}}
\caption{\label{fig:mich} 
Live installation at the gallery of Ludo Mich in Antwerp. There was no white board but different 
small pancartes with possible scenes (see in the middle of the picture). The game was displayed much larger against
the back wall.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap9/figs/look-into-box}}
\caption{\label{fig:lookintobox} 
Look into the Box installation at the Paris Museum of Modern Art. Left: Box with camera inside. A lens would make 
the eye look bigger for the camera. Right: Projection of the eye on a big screen.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.30\textwidth]{chap9/figs/eyes}
\includegraphics[width=.30\textwidth]{chap9/figs/identified-colours}}
\caption{\label{fig:eyes} 
Example interaction of the ``Look into the Box" art piece. 
Left: Top and pixilated eye of a spectator. Right: the main colours that 
were extracted from this image. Agents played language games to describe these eye colours.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap9/figs/ital-talk}
\includegraphics[width=.30\textwidth]{chap9/figs/ital2}}
\caption{\label{fig:spoleto} 
Left: My presentation at the Spoleto Science Festival. Right: Nicolas Neubauer debugging 
the Look into the Box installation.
 }
\end{figure}

The Look into the Box installation was also shown during the ``Intensive Science" exhibitions organised in October 
2006 on the occasion of the 10th anniversary of the Sony Computer Science Laboratory.\is{Intensive Science exhibition}
It was first shown in Paris at 
the exhibition space ``La Maison Rouge" near the Bastille in Paris, which featured science/art installations 
and collaborations involving members of the Sony Computer Science Laboratory, including work by Atau Tanaka, 
Francois Pachet in collaboration with Jazz pianist Albert van Veendaal, Peter Hanappe in collaboration with 
photographer Armin Linke, Frederic Kaplan in collaboration with the Design School of EPFL in Lausanne. 
The same exhibition traveled to Tokyo where it was shown in the Sony Explorascience museum from 22 December 2006 
to 12 February 2007 (see \figref{fig:intensive-science}). 
\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap9/figs/explorascience}}
\caption{\label{fig:intensive-science} 
Installation at the Tokyo Explorascience Museum. The box is in the left corner. The eye is projected on a 
big screen. A smaller screen shows the image captured by the camera, the colours recognised, and the words used by 
the agents.}
\end{figure}
\section{Conclusions}

>>>>

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.75\textwidth]{chap9/figs/cambridge-view.jpg}}
\caption{\label{fig:sideview} 
View through the street window inside the Kettle's Yard gallery in Cambridge. The cameras are located before 
the towers. The geometric figures are attached to the wall and explanations of the experiment are located on the right.}
\end{figure}

The N01SE exhibition was curated by Adam Lowe (see \figref{fig:lowe}) and Simon Schaffer. 
Adam Lowe is an artist and technologist. He is currently the director of Factum arte (Madrid) which is specialised 
in making life-like replicas of paintings, sculptures and archeological objects, using laser-scanners and 3d printers. 
Recent realisations include the sculptures of Giambattista Piranesi \citep{Lowe:2010}, the 
paintings of Caravaggio and the tomb of Tutankhamun. 
Simon Schaffer is a science historian, professor at the University of Cambridge at the department of History and Philosophy 
of Science. He wrote extensively about the historical developments in scientific research \citep{Schaffer:2011}
and animated television and radio programs, including {\itshape Light Fantastic} (BBC4). 
\begin{figure}[htbp]
  \centerline{\includegraphics[width=.70\textwidth]{chap9/figs/lowe.pdf}}
\caption{\label{fig:lowe} 
Adam Lowe (curator of N01SE) during installation of the Talking Heads experiment at the Wellcome Gallery in London. 
The backwall contained the figures. On the right wall, posters were displayed explaining the experiment.}
\end{figure}

@@

N01SE celebrates the essential excess from which information is drawn. It probes many different ways of seeing and 
being in the world. Chances are your own sense of order is already someone elses N01SE.}
\end{quotation}
%\begin{figure}[htbp]
%  \centerline{\includegraphics[width=.95\textwidth]{chap9/figs/latour.pdf}}
%\caption{\label{latour} 
%Luc Steels Discusses The N01SE exhibition with Bruno Latour who intensively followed the experiments already from the first series at the Laboratorium exhibition in Antwerp. His comments significantly improved the initial designs.}
%\end{figure}

@@

To play the game, the Speaker segments the scene and performs pattern recognition to extract features -- area, shape, colour, position -- about the segments selected. The Speaker then conceptualises the focus element ``topic" as distinct from those of other segments - be it the largest or the furthest to the left or the green one. Next the Speaker verbalises this conceptualisation using descriptive words selected from its lexicon. The Hearer works the other way around; it queries the words transmitted by the Speaker and applies the resultant meanings back to the scene to find what topic the Speaker intended.

@@

The installation consists of two computer-controlled robotic camera heads that capture images from ``scenes'' in front of them consisting of colored geometric shapes pasted on a magnetic whiteboard. The configurations on the board can be changed at any time, making the robots' world unpredictable and open.

@@

\begin{itemize} 
\item Two embodied grounded agents always see the situation from different points of view so that they capture different images. Consequently they may have divergent perceptions and hence great difficulty to arrive at a successful game. 
\item The word(s) transmitted may not be accurately produced or received. For example, one agent may produce `wabaku' but the other agent may hear `mabaku'. This introduces noise in the signal itself and hence possibly confusion among the agents. 
\item A scene can usually be conceptualised in many different ways, so that there is seldom certainty among the agents whether they share the same meaning. This causes great difficulties in learning the meaning of unknown words. 
\item The lexicons and conceptual repertoires are never exactly the same as each agent develops them autonomously. This generates in additional sources of confusion.
\end{itemize}

@@

Another factor in language evolution is due to changes in the environment. Thus two alternative meanings for a word may be compatible for a while but are then disambiguated when a series of scenes arises in which the two meanings are no longer both applicable. For example, all objects may be both green and small and therefore there may be a word `sesubipu' which may mean both, until a clear situation arises where a green object is no longer the smallest and a misunderstanding arises.

Semiotic evolution is continually present. Different meanings for a particular word will emerge over over a large number of language games. During specific periods, different words dominate. The word `droite' (originally introduced by a French speaker) gains the dominant meaning `to the right', then shifts to `at the bottom', and then to `very much to the right'. Particularly the words introduced by humans have a tendency to undergo this kind of strong evolution because human users do not know which meanings their agents employ.

@@

Each agent has been given a mechanism to `grow' new distinctions by expanding discrimination trees. Each tree discretises one sensory dimension. For example, there is a tree for the area of a segment (scale with respect to the image) which divides the range of possible values into two discrete regions, which would be named in English `small' and `large'. Other trees focus on position (left versus right or top versus bottom), shape (rectangular or oval), color, etc. Trees can go as deep as necessary to carve out smaller and smaller subregions of a continuous space.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.60\textwidth]{chap9/figs/NOISE-programmaboek.jpg}}
\caption{\label{fig:program} 
Catalogue cover and poster of the N01SE exhibition at the Wellcome Gallery in London.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.55\textwidth]{chap9/figs/london-heads.pdf} 
\includegraphics[width=.44\textwidth]{chap9/figs/screen.pdf}}
\caption{\label{fig:london-heads} 
Installation at the Wellcome Gallery in London. Left: Talking Heads cameras oriented towards the wall on which 
geometric figures were pasted. Right: Projection of interaction during ongoing experiment. A game just failed and 
the speaker says ``no''.} 
\end{figure}

@@

\noindent
The installation consists of two robotic heads. These are steerable cameras controlled by a computer that hosts the 
architecture and knowledge state of each agent. The robots capture images from scenes in front of them. The scenes consist 
of coloured geometrical shapes pasted on a magnetic white board. The configuration on the board can be changed at any 
time, making the robots' world unpredictable and open. The robot infrastructure is connected to the Internet, so 
that an agent may dematerialise from a body and travel over the internet to another body in which it can re-materialise.
Two robotic structures will be active as part of the N01SE exhibition: one at Kettles'Yard in Cambridge and another 
one at the Wellcome Trust Two10 Gallery in London.

@@

\noindent
What is most remarkable about this experiment is that the robotic agents do not come with 
pre-programmed ways of conceptualising reality but have to develop their own concepts. Each agent has been 
given a mechanism to `grow' new distinctions by expanding discrimination trees. Each tree discretises one sensory 
dimension. For example, there is a tree for the area of a segment which divides the range of possible values
into two discrete regions, which would be named in English `small' and `large'. Other trees focus on position 
(left versus right or top versus bottom), shape (rectangular or oval), colour, etc.

@@

\noindent
The creation of a shared language by a group of autonomous distributed agents is extraordinarily difficult 
because there are many sources of noise, in the form of disturbances that cause incoherence between the agents: 
\begin{itemize} 
\item Two embodied grounded agents always see the situation from different points of view so that they capture 
different images. Consequently they may have divergent perceptions and hence great difficulty to arrive at a 
successful game. 
\item The word(s) transmitted may not be accurately produced or received. For example, one agent may produce 
'wabaku' but the other agent may hear `mabakau'. This introduces noise in the signal itself and hence 
possibly confusion among the agents. 
\item A scene can usually be conceptualised in many different ways, so that there is seldom certainty among 
the agents whether they share the same meaning. This causes great difficulties in learning 
the meaning of unknown words. 
\item The lexicons and conceptual repertoires are never exactly the same due to the fact that each agent
develops them autonomously. This brings in additional sources of confusion. 
\end{itemize}
Any theory claiming to explain the origins of word meaning must confront the handling of 
noise head on.

@@

\noindent
Noise plays yet another role, namely as a motor of language evolution. Indeed natural lexicons evolve - 
even if there are already perfectly good words in a language. Noise on the word form causes changes 
in the form which propagate in the remainder of the population. Misunderstandings may destabilise a
word and cause its meaning to shift.

@@

Many aspects of this episode are remarkable, not in the least that those responsable for an 
exhibition (i.c. Denna Jones (i.e. phlox) and ``fish'') were bent on creating a wave of negative reactions and 
destructions against one of the exhibition pieces entrusted to them. Apparently this behaviour was triggered through
a conflict 
with Adam Lowe, curator of the N01SE exhibition, but the team behind the Talking Heads experiment had nothing to do 
with this. Another remarkable fact is that the hacker group 
became aggressive as soon as they felt they were no longer able to have control the way they wanted to, specifically 
to introduce disrespectful language or to subvert agent scripts to circumvent the central scheduler so that they 
could send their agents in priority to the sites of their choosing.
This style of behaviour is a personality trait which is commonly recognised as characteristic for 
hackers, including by members of the hacker community themselves:\footnote{
Raymond, E. (2013) The new hackers dictionary. Available as http://www.catb.org/jargon/html/index.html.}

\begin{quotation}
``Hackers have relatively little ability to identify emotionally with other people. ... Unsurprisingly, hackers also tend towards self-absorption, intellectual arrogance, and impatience with people and tasks perceived to be wasting their time. Because of their passionate embrace of (what they consider to be) the Right Thing, hackers can be unfortunately intolerant and bigoted on technical issues, in marked contrast to their general spirit of camaraderie and tolerance of alternative viewpoints otherwise. ... As a result of all the above traits, many hackers have difficulty maintaining stable relationships. At worst, they can produce the classic geek: withdrawn, relationally incompetent, sexually frustrated, and desperately unhappy when not submerged in his or her craft."
\end{quotation}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap9/figs/th4.jpg}}
\caption{\label{fig:palais} 
Talking Heads installation at the Palais de la D\'{e}couverte. It featured new fancy structures that 
made the installation more attractive visually. Daily explanations were given to visitors 
by the staff of the Palais de la D\'{e}couverte.} 
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.60\textwidth]{chap9/figs/road.pdf}}
\caption{\label{fig:road} 
Portable installation with two pan-tilt cameras and a portable computer that was able to run the TH software.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.90\textwidth]{chap9/figs/Lausanne.pdf}}
\caption{\label{fig:lausanne} 
Live installation and demonstration of the Talking Heads experiment at the European Conference on Artificial 
Life in 1999.}
\end{figure}

2. The {\itshape Neuer Aachener Kunstverein} in Aachen (Germany) in collaboration with the 
RWTH (the technical university of Aachen) organised a general exhibition on modelling called 
Modell-Modell. Within this framework, artist Anne-Mie van Kerckhoven invited me to cooperate in a laboratory on 
language and colour called ``Cyberlabor Chromosophy''. As part of this laboratory, the portable Talking Heads 
experiment was installed and ran from 5 May to 16 June 2000. It featured not only the portable Talking Heads, 
which was demonstrated live, but also posters, talks about the project and additional art works.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap9/figs/mich-gallery.pdf}}
\caption{\label{fig:mich} 
Live installation at the gallery of Ludo Mich in Antwerp. There was no white board but different 
small pancartes with possible scenes (see in the middle of the picture). The game was displayed much larger against
the back wall.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap9/figs/look-into-box.pdf}}
\caption{\label{fig:lookintobox} 
Look into the Box installation at the Paris Museum of Modern Art. Left: Box with camera inside. A lens would make 
the eye look bigger for the camera. Right: Projection of the eye on a big screen.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.30\textwidth]{chap9/figs/eyes.pdf}
\includegraphics[width=.30\textwidth]{chap9/figs/identified-colors.pdf}}
\caption{\label{fig:eyes} 
Example interaction of the ``Look into the Box" art piece. 
Left: Top and pixilated eye of a spectator. Right: the main colours that 
were extracted from this image. Agents played language games to describe these eye colours.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap9/figs/ital-talk.pdf}
\includegraphics[width=.30\textwidth]{chap9/figs/ital2.pdf}}
\caption{\label{fig:spoleto} 
Left: My presentation at the Spoleto Science Festival. Right: Nicolas Neubauer debugging 
the Look into the Box installation.
 }
\end{figure}

The Look into the Box installation was also shown during the ``Intensive Science" exhibitions organised in October 
2006 on the occasion of the 10th anniversary of the Sony Computer Science Laboratory.\is{Intensive Science exhibition}
It was first shown in Paris at 
the exhibition space ``La Maison Rouge" near the Bastille in Paris, which featured science/art installations 
and collaborations involving members of the Sony Computer Science Laboratory, including work by Atau Tanaka, 
Francois Pachet in collaboration with Jazz pianist Albert van Veendaal, Peter Hanappe in collaboration with 
photographer Armin Linke, Frederic Kaplan in collaboration with the Design School of EPFL in Lausanne. 
The same exhibition traveled to Tokyo where it was shown in the Sony Explorascience museum from 22 December 2006 
to 12 February 2007 (see \figref{fig:intensive-science}). 
\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap9/figs/explorascience.pdf}}
\caption{\label{fig:intensive-science} 
Installation at the Tokyo Explorascience Museum. The box is in the left corner. The eye is projected on a 
big screen. A smaller screen shows the image captured by the camera, the colours recognised, and the words used by 
the agents.}
\end{figure}
\section{Conclusions}
